Verification of Stencil Computations with Spatial Specifications

### 2016-07-07 paper.tex
edit

\emph{Stencils} are a ubiquitous programming pattern, common in
scientific and numerical computing applications. Informally, a stencil
computation outputs an array, where the value at each index $i$ of
this array is calculated from a \emph{neighbourhood} of values around $i$ in
some input array(s), \eg{}, the Game of Life, convolutions in image
processing, approximations to differential equations. For example, the
following computes the one-dimensional discrete Laplace
transform (an approximation to a derivative) in Fortran:
%
\begin{minted}{fortran}
do iter = 0, itermax
   do i = 1, (n-1)
      b(i) = a(i-1) - 2*a(i) + a(i+1)
   end do
   a = b
end do
\end{minted}
%
Line 3 is the core of the stencil computation, calculating
the value at \texttt{b(i)} from a neighbourhood of elements about
\texttt{a(i)}. Line 5 swaps
\texttt{a} and \texttt{b} between iterations, where \texttt{b} becomes the
input for the next iteration. In Fortran, parentheses \texttt{( )} are used
for array subscripts rather than the more familiar bracket syntax \texttt{[ ]}.

%This stencil computation exhibits statically decideable
%spatial and temporal relationships between $\texttt{a}$ and
%$\texttt{b}$.
The shape of the data access pattern on line 3 determines other aspects of the
program and its efficient implementation: how much ``boundary'' is
needed for the array, the most cache-efficient layout in memory,
and the partitioning shape for parallel implementations.
In this example, the access pattern is simple, static, and easily
understood. More complex stencil computations
are much more prone to error from simple lexical mistakes and
harder to understand. For example,
Figure~\ref{ref:navier-stokes-fragment} shows three lines from a
Navier-Stokes fluid simulator in which two arrays are read from with
different data access patterns, across two dimensions. The interaction
is much harder to understand, with the potential for the developer to
accidentally introduce an error via simple textual mistakes, for
example writing $\texttt{(i-1,j)}$ instead of $\texttt{(i+1,j)}$.

@@

\item we provide an implementation of our approach as an extension to
  CamFort, an open-source program analysis tool for Fortran programs.
  We use this for a quantitative evaluation of our approach over a corpus of
  scientific applications, totalling roughly one million lines of
  code;

>>>>

\emph{Stencils} are a ubiquitous programming pattern, common in
scientific and numerical computing applications. Informally, a stencil
computation outputs an array, where the value at each index $i$ of
this array is calculated from a \emph{neighbourhood} of values around $i$ in
some input array(s), \eg{}, the Game of Life, convolutions in image
processing, approximations to differential equations. For example, the
following computes the one-dimensional discrete Laplace
transform (an approximation to a derivative) in Fortran:
%
\begin{minted}{fortran}
do iter = 0, itermax
   do i = 1, (n-1)
      b(i) = a(i-1) - 2*a(i) + a(i+1)
   end do
   a = b
end do
\end{minted}
%
Line 3 is the core of the stencil computation, calculating
values at \texttt{b(i)} from a neighbourhood of elements about
\texttt{a(i)}. Line 5 swaps
\texttt{a} and \texttt{b} between iterations, where \texttt{b} becomes the
input for the next iteration. %In Fortran, parentheses \texttt{( )} are used
%for array subscripts rather than the more familiar bracket syntax \texttt{[ ]}.
%
%This stencil computation exhibits statically decideable
%spatial and temporal relationships between $\texttt{a}$ and
%$\texttt{b}$.
The shape of the data access pattern on line 3 determines other aspects of the
program and its efficient implementation: how much ``boundary'' is
needed for the array, the most cache-efficient layout in memory,
and the partitioning shape for parallel implementations.

In this example, the access pattern is simple, static, and easily
understood. More complex stencil computations
are much more prone to error from simple lexical mistakes and
harder to understand. For example,
Figure~\ref{ref:navier-stokes-fragment} shows three lines from a
Navier-Stokes fluid simulator in which two arrays are read from with
different data access patterns, across two dimensions. The interaction
is much harder to understand, with the potential for the developer to
accidentally introduce an error via simple textual mistakes, for
example writing $\texttt{(i-1,j)}$ instead of $\texttt{(i+1,j)}$.

@@

\item we provide an implementation of our approach as an extension to
  CamFort, an open-source program analysis tool for Fortran programs.
